---
title: "Updated Analysis"
author: "Ian Scarff"
date: "September 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prep Data

Load in data. This new file includes a variable for month.

```{r}
FB <- read.csv("FacebookDataReduced.csv")

### Be sure that NAME is a character variable
FB$name <- as.character(FB$name)

### Remove any null
FB <- FB[FB$name != "null",]
```

# Summary Graphs

Load *plotly* package.

```{r, message=FALSE,warning=FALSE}
library(plotly)
```

### Pie Charts
```{r}
plot_ly(data = FB, labels = ~Source, type = 'pie') %>%
  layout(title = 'Data by News Source',
         xaxis = list(showgrid = F, zeroline = F, showticklabels = F),
         font = list(size = 24))
```


```{r}
plot_ly(data = FB, labels = ~year, type = 'pie', textposition = 'inside',
        textinfo = 'label+percent', text = ~paste('Year: ', year),
        showlegend = F) %>%
  layout(title = 'Data by Year',
         xaxis = list(showgrid = F, zeroline = F, showticklabels = F),
         font = list(size = 24))
```


### Bar Charts

Create a vector of years for later use.

```{r}
years <- c(2012:2016)
```

The bar chart below shows the total number of stories for each new outlet, across each year.

```{r}
### Create a data frame to hold total story values
TotStories <- as.data.frame(matrix(data = NA, nrow = 15, ncol = 6))

### Set column names to correspond to News and Years
colnames(TotStories) <- c("News",paste("Y",years,sep=""))

### Set the first column to the news outlet
TotStories[,1] <- levels(FB$Source)

### This for loop will calculate the total stories and insert them into the data frame
### For each news source
for (i in 1:length(levels(FB$Source))){
  
  ### For each year
  for (j in 1:length(years)){
    
    ### Check to see if the news outlet has data in the selected year
    if (any((subset(FB, (Source == levels(FB$Source)[i]), select = year)) == years[j])){
      
      ### If it does, take a subset with the selected year and news outlet, grab a random variable variables,
      ### find out how many rows are in the subset, and add it to the data frame
      TotStories[i,j+1] <- nrow(subset(FB, (Source == levels(FB$Source)[i] & year == years[j]), 
                                    select = c(shares_count)))
    }
    
    ### If there is no data for a news outlet in the selected year
    else{
      
      ### Set total reactions to zero and add it to the data frame
      TotStories[i,j+1] <- 0
    }
  }
}


### Plot bar chart
plot_ly(TotStories, x = ~News, y = ~Y2012, type = 'bar', name = '2012') %>%
  add_trace(y = ~Y2013, name = '2013') %>%
  add_trace(y = ~Y2014, name = '2014') %>%
  add_trace(y = ~Y2015, name = '2015') %>%
  add_trace(y = ~Y2016, name = '2016') %>%
  layout(yaxis = list(title = '# of Stories'), barmode = 'group')

```




The bar chart below shows the total number of reactions (Likes + Love + Sad + Thankful + Angry) for each new outlet, across each year.

```{r}

### Create a data frame to hold total reaction values
TotReact <- as.data.frame(matrix(data = NA, nrow = 15, ncol = 6))

### Set column names to correspond to News and Years
colnames(TotReact) <- c("News",paste("Y",years,sep=""))

### Set the first column to the news outlets
TotReact[,1] <- levels(FB$Source)

### This for loop will calculate the total reactions and insert them into the data frame
### For each news source
for (i in 1:length(levels(FB$Source))){
  
  ### For each year
  for (j in 1:length(years)){
    
    ### Check to see if the news outlet has data in the selected year
    if (any((subset(FB, (Source == levels(FB$Source)[i]), select = year)) == years[j])){
      
      ### If it does, take a subset with the selected year and news outlet, grab the reaction variables, sum them up,
      ### and add it to the data frame
      TotReact[i,j+1] <- sum(subset(FB, (Source == levels(FB$Source)[i] & year == years[j]), 
                                  select = 
                          c(likes_count,love_count,wow_count,haha_count,sad_count,
                                      thankful_count,
                                      angry_count)))
    }
    
    ### If there is no data for a news outlet in the selected year
    else{
      
      ### Set total reactions to zero and add it to the data frame
      TotReact[i,j+1] <- 0
    }
  }
}


### Plot the bar chart
plot_ly(TotReact, x = ~News, y = ~Y2012, type = 'bar', name = '2012') %>%
  add_trace(y = ~Y2013, name = '2013') %>%
  add_trace(y = ~Y2014, name = '2014') %>%
  add_trace(y = ~Y2015, name = '2015') %>%
  add_trace(y = ~Y2016, name = '2016') %>%
  layout(yaxis = list(title = '# of Reactions'), barmode = 'group')

```


The bar chart below shows the total number of comments for each new outlet, across each year.


```{r}
### Create a data frame to hold total comment values
TotComm <- as.data.frame(matrix(data = NA, nrow = 15, ncol = 6))

### Set column names to correspond to News and Years
colnames(TotComm) <- c("News",paste("Y",years,sep=""))

### Set the first column to the news outlets
TotComm[,1] <- levels(FB$Source)

### This for loop will calculate the total comments and insert them into the data frame
### For each news source
for (i in 1:length(levels(FB$Source))){
  
  ### For each year
  for (j in 1:length(years)){
    
    ### Check to see if the news outlet has data in the selected year
    if (any((subset(FB, (Source == levels(FB$Source)[i]), select = year)) == years[j])){
      
      ### If it does, take a subset with the selected year and news outlet, grab the comment variable, sum it up,
      ### and add it to the data frame
      TotComm[i,j+1] <- sum(subset(FB, (Source == levels(FB$Source)[i] & year == years[j]), 
                                    select = c(comments_count)))
    }
    
    ### If there is no data for a news outlet in the selected year
    else{
      
      ### Set total comments to zero and add it to the data frame
      TotComm[i,j+1] <- 0
    }
  }
}


### Plot the bar chart
plot_ly(TotComm, x = ~News, y = ~Y2012, type = 'bar', name = '2012') %>%
  add_trace(y = ~Y2013, name = '2013') %>%
  add_trace(y = ~Y2014, name = '2014') %>%
  add_trace(y = ~Y2015, name = '2015') %>%
  add_trace(y = ~Y2016, name = '2016') %>%
  layout(yaxis = list(title = '# of Comments'), barmode = 'group')
```

The bar chart below shows the total number of shares for each new outlet, across each year.

```{r}
### Create a data frame to hold total share values
TotShare <- as.data.frame(matrix(data = NA, nrow = 15, ncol = 6))

### Set column names to correspond to News and Years
colnames(TotShare) <- c("News",paste("Y",years,sep=""))

### Set the first column to the news outlets
TotShare[,1] <- levels(FB$Source)

### This for loop will calculate the total shares and insert them into the data frame
### For each news source
for (i in 1:length(levels(FB$Source))){
  
  ### For each year
  for (j in 1:length(years)){
    
    ### Check to see if the news outlet has data in the selected year
    if (any((subset(FB, (Source == levels(FB$Source)[i]), select = year)) == years[j])){
      
      ### If it does, take a subset with the selected year and news outlet, grab the shares variable, sum it up,
      ### and add it to the data frame
      TotShare[i,j+1] <- sum(subset(FB, (Source == levels(FB$Source)[i] & year == years[j]), 
                                   select = c(shares_count)))
    }
    
    ### If there is no data for a news outlet in the selected year
    else{
      
      ### Set total comments to zero and add it to the data frame
      TotShare[i,j+1] <- 0
    }
  }
}


### Plot the bar chart
plot_ly(TotShare, x = ~News, y = ~Y2012, type = 'bar', name = '2012') %>%
  add_trace(y = ~Y2013, name = '2013') %>%
  add_trace(y = ~Y2014, name = '2014') %>%
  add_trace(y = ~Y2015, name = '2015') %>%
  add_trace(y = ~Y2016, name = '2016') %>%
  layout(yaxis = list(title = '# of Shares'), barmode = 'group')
```


### Box Plots


Since the data is highly right scewed, a log transformation was used.

* **NOTE:** In statistics, whenever a log transformations is mentioned/used, it is **always** a *natural log* transformation, not log base 10.

The box plot below shows the log transformation of likes for each news outlet.

* Because the minimum value is 0, the data was shifted by 1. (log(0) = -Infinity)

```{r, warning=FALSE}
plot_ly(data = FB, y = ~log(likes_count + 1), color = ~Source, type = 'box') %>%
  layout(yaxis = list(title = "Log of (Likes + 1)"))
```

The box plot below shows the log transformation of comments for each news outlet.

* Because the minimum value is 0, the data was shifted by 1. (log(0) = -Infinity)

```{r,warning=FALSE}
plot_ly(data = FB, y = ~log(comments_count + 1), color = ~Source, type = 'box') %>%
  layout(yaxis = list(title = "Log of (Comments + 1)"))
```

The box plot below shows the log transformation of comments for each news outlet.

* Because the minimum value is 0, the data was shifted by 1. (log(0) = -Infinity)

```{r,warning=FALSE}
plot_ly(data = FB, y = ~log(shares_count + 1), color = ~Source, type = 'box') %>%
  layout(yaxis = list(title = "Log of (Shares + 1)"))
```


# Text Mining

Load package 
```{r, message=FALSE,warning=FALSE}
library(tm)
```


The following function will take in a data frame (all varaibles included), and conduct the text mining procedure. It will return a data frame with words and their frequencies.

```{r}
textFreq <- function(Data){
  
  names_text <- paste(Data[3], collapse = " ")
  
  ### Source and corpus
  names_source <- VectorSource(names_text)
  names_corpus <- Corpus(names_source)
  
  ### Clean
  names_corpus <- tm_map(names_corpus, content_transformer(tolower))
  names_corpus <- tm_map(names_corpus, removePunctuation)
  names_corpus <- tm_map(names_corpus, stripWhitespace)
  names_corpus <- tm_map(names_corpus, removeWords, stopwords("english"))
  
  ### making a document-term matrix
  
  names_dtm <- DocumentTermMatrix(names_corpus)
  names_dtm <- as.matrix(names_dtm)
  
  
  ### Finding the most frequent terms
  names_frequency <- colSums(names_dtm)
  names_frequency <- sort(names_frequency, decreasing = T)
  
  words <- names(names_frequency)

  return(data.frame(words,names_frequency))
}
```


# Analysis


First, look at the most popular words in stories that have a number of likes greater than or equal to the median number of likes.

```{r,message=FALSE,warning=FALSE,error=FALSE}
greaterThanMedian1 <- textFreq(FB[FB$likes_count >= median(FB$likes_count),])
```

Second, look at the most popular words in stories that have a number of comments greater than or equal to the median number of comments.

```{r,message=FALSE,warning=FALSE,error=FALSE}
greaterThanMedian2 <- textFreq(FB[FB$comments_count >= median(FB$comments_count),])
```

### Word Clouds

For some reason, the two most popular words are *timeline* and *photo*. These words aren't interesting. Excluding those, the following word clouds show the top 100 words across the whole dataset.

* *NOTE*: The top 30 words are in red.
```{r, message=FALSE,warning=FALSE}
library(wordcloud)
```



```{r}
wordcloud(greaterThanMedian1$words[3:102], 
          greaterThanMedian1$names_frequency[3:102], random.order = F,
          colors = c((rep("red", times = 30)), rep("black", times = 70)),
          ordered.colors = T,)
          
text(x=0.5, y=0.98, "Top 100 Words where Story Likes are >= Median", cex = 1)
```


```{r}
wordcloud(greaterThanMedian2$words[3:102], 
          greaterThanMedian2$names_frequency[3:102], random.order = F,
          colors = c((rep("red", times = 30)), rep("black", times = 70)),
          ordered.colors = T)
text(x=0.5, y=0.98, "Top 100 Words where Story Comments are >= Median", cex = 1)
```



# What To Do Next

Examine stories by year, month, news outlet, and combinations of all. Expand the analysis to other reactions.


